---
title: "Preparing Benchmark Data"
author: "Niklas Rindtorff"
date: "2022-12-18"
output: md_document
---

# Preparing Benchmark Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(here)
library(readxl)
```

The D3R challenge is a competition for models of small-molecule protein interaction. The challenge consists of a pose prediction as well as a [[affinity ranking]] task.

We focus on the affinity ranking task and curate an oppinionated benchmark dataset that spans all data in which more than 100 ligands are available for a given protein target.

In the spirit of zero-shot predictions, we store the FASTA file of the protein target as well as a reference PDB file that is publicly available from the PDB.

```{r}
# formatting varies widely between the data sets, as they are from different years
bace <- read_csv(here("data/external/d3r_benchmark/BACE1/BACE_score_compounds_D3R_GC4_answers.csv")) %>% dplyr::select(ligand_id = Cmpd_ID,
              ligand = SMILES,
              affinity = Affinity) %>% 
  mutate(gene = "BACE1")
cats <- read_csv(here("data/external/d3r_benchmark/CatS/CatS_score_compounds_D3R_GC4_answers.csv")) %>% dplyr::select(ligand_id = Cmpd_ID,
              ligand = SMILES,
              affinity = Affinity) %>% 
  mutate(gene = "CTSS")
jak2 <- read_csv(here("data/external/d3r_benchmark/JAK2_SC2/JAK2_SC2_score_compounds_D3R_GC3.csv")) %>%
  dplyr::select(ligand_id = Ligand_ID,
              ligand = Smiles,
              affinity = `Kd(uM)`) %>%
  mutate(gene = "JAK2")
p38a <- read_csv(here("data/external/d3r_benchmark/p38a/p38a_score_compounds_D3R_GC3.csv")) %>%
  dplyr::select(ligand_id = Ligand_ID,
              ligand = Smiles,
              affinity = `Kd(uM)`) %>%
  mutate(gene = "MAPK14")

cdk2 <- read_excel(here("data/external/d3r_benchmark/CDK2_Binding_Data_Corrected_2016AUG18.xlsx.xlsx")) %>%
  dplyr::select(ligand_id = Compound_ID,
              ligand = SMILES,
              affinity = `Experimental Result, Average Value`,
              pdb = `PDB ID`) %>%
  mutate(gene = "CDK2") %>% 
  mutate(unit = "nM") %>% 
  drop_na()

```

```{r}
#todo import all datasets, check methods and units of affinity
```
