---
title: "SIRT3 FDA affinity ranking"
author: "Niklas Rindtorff"
date: "2022-12-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```


# Library creation

## downloading library
we start out by installing a set of FDA approved compounds from an online source. This is list is not fully up to date, but should be sufficiently relevant to serve as a useful set to generate a reference affinity ranking that we can use to compare candidate compounds against. 

```{r}
system("wget http://www.cureffi.org/wp-content/uploads/2013/10/drugs.txt; mv drugs.txt data/external/fda_affinity_ranking/fda.txt")
```
we next subsample the set of FDA approved drugs to small molecules with an available SMILES

```{r}
fda <- read_delim(here("data/external/fda_affinity_ranking/fda.txt"), 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% 
  drop_na()
```
## subsampling library and exporting

```{r}
set.seed(24234)
fda_sirt3 = fda %>% 
  sample_n(498) %>%
  dplyr::select(ligand_id = generic_name, ligand = smiles) %>%
  rbind(c("5514", "O=C(COC1=CC=C(C[C@H](OCC)C(N(C)OC)=O)C=C1)C2=CC(OC)=CC=C2"), .) %>% 
  rbind(c("NAD-Ribose", "c1nc(c2c(n1)n(cn2)[C@H]3[C@@H]([C@@H]([C@H](O3)COP(=O)(O)OP(=O)(O)OC[C@@H]4[C@H]([C@H]([C@H](O4)O)O)O)O)O)N"), .) %>% 
   # the path below expects the sirt3 repo to be cloned to the root dir
  mutate(protein_path = "/tmp/pdb/4bn4_protein_stripped_no_ions.pdb") %>%
  # only writing 5 for testing
  head(10) %>%
  write_csv(here("data/processed/affinity_ranking/SIRT3/input_protein_ligand.csv"))
  
```

# Docking analysis 

```{r}
library(readr)
fda_sirt3_ranking <- read_delim(here("data/processed/affinity_ranking/test/diffdock_221222 SIRT3-5/df_diffdock_results.tsv"), 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% 
  mutate(replicate = 2) %>% 
  rbind(read_delim(here("data/processed/affinity_ranking/test/diffdock_221220 SIRT3-5/df_diffdock_results.tsv"), 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% 
  mutate(replicate = 1))
```

```{r}
fda_sirt3_ranking %>% 
  dplyr::select(ligand = smiles, everything()) %>%
  left_join(fda_sirt3) %>%
  ggplot(aes(diffdock_confidence, smina_scored_affinity, color = ligand_id)) + 
    geom_point()
```

